Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-31 manuscript/manuscript.Rmd
adds new results from upping the perturbation percentages for climate

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may difficult resolve in statistical models based on population-level data where demographic processes are not identifiable. If population-level data cannot detect important impacts of climate drivers on populations, then population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on aggregated, population-level data can predict climate change impacts as wells as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. [THE REST OF THIS PARAGRAPH READS TOO MUCH LIKE AN ABSTRACT. CUT OR REDUCE TO ONE SENTENCE? e.g. "individual levels models were clearly better, but, unfortunately, still too uncertain to inform management"] We found that population models based on detailed demographic data are more accurate and precise than models based on aggregated data. Our results suggest the popuation-level model is less accurate and precise because important demographic climate signals go undetected. For these species at this location, detailed demographic data appears necessary to make accurate forecasts. A worrying caveat to our work is that forecasts from both models were very uncertain when we considered full process and parameter uncertainty. Even 14 years worth of demographic data may not be sufficient to make useful forecasts.

@@

We modeled growth as Gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \sigma_{j})
\end{align}

?Is sigma_j a constant, or does it depend on other parameters or state variables?

where $x$ is log genet size and all other parameters are as described for the survival regression.

@@

\begin{align}
x_{jq,t+1} &= \gamma^{P}_{j,t} + \phi^{P}_{jQ} + \beta^{P}_{j,t}x_{jq,t} + \theta^{P}_{jk}C_{k,t} \\
y^{P}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) \text{T}[0,1]
\end{align}

@@

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual level data [cite Ellner and Rees 2006] and a quadrat-based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat level data. We describe each in turn.

#### Integral projection model
We use an environmentally stochastic IPM [@Rees2009] that includes the random year effects and the climate covariates from the vital rate statistical models. However, for some simulations, we ignore the random year effects so that only the climate effects drive interannual variation \todo{Add spatial structure mention. Maybe you can specify when we use random yrs effects, when not. Just above line 191 I found myself wondering when you do use them }. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size _v_ at time $t+1$ is

@@

@@

which means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species. Our stastical model for recruitment (_R_, described above) returns the number of new recruit produced per quadrat. Following previous work [@Adler2012; @Chu2015], we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM. \todo{Add spatial structure mention.}

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eqs. 8-9). We use the same approach for drawing parameter values as described for the IPM. After drawing the appropriate parameter set, we calculate the mean response (population cover at *t+1* = $x_{t+1}$) according to Eq. 8. We then make a random draw from a [0,1] truncated lognormal distribution with mean equal to $x_{t+1}$ from Eq. 8 and the variance estimate from the fitted model. We can then project the model forward by drawing a new parameter set (unique to climate year and MCMC iteration) at each timestep.

@@

### Testing sensitivity to climate covariates
Our main goal in this paper \todo{is this our main goal? it's not quite the same as making good predictions. It is a goal...} is to see if models based on aggregated data are as sensitive to climate covariates as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

Our expectation is that the IPM will produce more accurate and precise forecasts. This could be due to 1) the smaller  sample size of the quadrat level data sets compared to the individual level data sets, leading to larger parameter uncertainty for the QBM, or 2) the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty we refit the QBM statistical model (Eqs. 8-9) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptlag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress sample size against parameter uncertainty.

@@

Results
-------
### Comparison of forecast models
The IPM had significantly lower overall error (MAE, mean absolute error) for three species (*B. gracilis*, *H. comata*, *P. smithii*; Table 1). \todo{Perhaps mention the PASM recruitment*climate uncertainty here? to explain why this species is different?} In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

@@

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6). \todo{I thought you were going to show the no-parameter-uncertainty forecasts in appendix?.} As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to a very high instrinsic rate of recruitment combined with uncertainty in climate coefficients which lead to high recruitment boom years and subsequent busts when young plants suffer high mortality (Appendx C). When we included model error and parameter uncertainty, forecast changes in proportional cover always spanned a wide range of negative to positive values. In other words, neither model could predict whether a climate perturbation would increase or decrease equilibrium population size.

@@

```{r comment_out, echo=FALSE, eval=FALSE}
Equilibrium cover from both models was sensitive to climate (Fig. 3a-d). The IPM projected percent changes in equilibrium cover from `r round(min(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)`% for *B. gracilis*, `r round(min(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)`% for *H. comata*, `r round(min(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)`% for *P. smithii*, and `r round(min(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)`% for *P. secunda*. The QBM projected opposite and greater percent changes in equilibrium cover for *B. gracilis* (`r round(min(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)`%) and *H. comata* (`r round(min(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)`%; Fig. 3a-b). For *P. smithii*, the QBM projected opposite changes in equilibrium cover than the IPM, but of similar magnitude (`r round(min(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)`%; Fig. 3c). *P. secunda* was the only species that the IPM and QBM made projections of the same sign and somewhat similar magnitude (`r round(min(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)`%; Fig. 3d). 
```

@@

Our comparison of a traditional, demographic population model (the IPM) with a model inspired by density-structured models (the QBM) showed that the IPM outperformed the QBM: the IPM was more accurate and precise than the QBM in out-of-sample cross validation (Table 1). The superiority of the IPM could reflect either differences in sample size or the effect of averaging over unique effects of climate on each individual-level vital rate. Although increasing sample size of quadrat percent-cover observations would be easy to do in the field, we little evidence that it would lead to higher precision of climate coefficient estimates (Fig. 4). We did, however, find evidence that the QBM statistical model failed to identify climate dependence for some vital rates (Fig. 5). For no species were climate effects from the QBM strongly correlated with all three vital rates (Fig. 5). @Freckleton2011 acknowledge that averaging over complex stage dependence will lead to poorly specified models. This is analagous to our situation, but instead of averaging over complex life histories, we are averaging over complex climate dependence.

@@

An important, but unexpected, result of our analysis was the great uncertainty in forecasts, even for our best model. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model, usually allowing for some variability around the determinstic process (e.g., @Battin2007; @Jenouvrier2009; @Adler2012). If we follow tradition and calculate the mean response to climate perturbation with only model error and interannual variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). It would be tempting to interpret this inconsistency as further evidence for the superiority of the IPM. However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6), though the IPM projections are generally more precise (consistent with our cross-validation results). The real story is that both models produce highly uncertain forecasts. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. \todo{I think we should look at larger perturbations. How large a perturbation do we need to get a forecast that doesn't span 0?}

@@

Our results also offer a cautionary tale because forecast uncertainty was large for both model types.Even with 14 years of detailed demographic data and sophisticated modeling techniques, we failed to produce forecasts with any level of acceptable uncertainty. Uncertainty of climate change related forecasts can be reduced by (1) longer time series and (2) climate covariates that match the scale of inference (e.g., plot rather than landscape level climate/weather metrics).

@@

```{r figure_6, dependson="plot-options", fig.cap="Equilibrium cover and 90% quantiles around the mean prediction when model error and parameter uncertainty are propogated through the simulation phase. Climate simulations are as in Figure 3.", fig.width=4.5, fig.height=7, cache=FALSE, results='hide'}
####
##  Vary parameter runs
####
setwd("~/Repos/MicroMesoForecast/analysis/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/climchange_varyparams/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]

>>>>

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. If population-level data cannot detect important impacts of climate drivers on populations, then population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on aggregated, population-level data can predict climate change impacts as wells as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. Individual level models were clearly better, but, unfortunately, still too uncertain to inform management decisions.

@@

We modeled growth as a Gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \varepsilon_{ij,t})
\end{align}

where $x$ is log genet size and all other parameters are as described for the survival regression. We capture non-constant error variance in growth by modeling the variance around the growth regression ($\varepsilon$) as a nonlinear function of predicted genet size:

\begin{align}
\varepsilon_{ij,t} = a e^{b x_{ijQ,t+1}}
\end{align}

@@

\begin{align}
x_{jq,t+1} &= \gamma^{P}_{j,t} + \phi^{P}_{jQ} + \beta^{P}_{j,t}x_{jq,t} + \theta^{P}_{jk}C_{k,t} \\
y^{P}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau_{j}) \text{T}[0,1]
\end{align}

@@

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual level data [@Ellner2006] and a quadrat-based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat level data. We describe each in turn.

#### Integral projection model
We use an environmentally stochastic IPM [@Rees2009] that includes the climate covariates from the vital rate statistical models. In all simulations we ignore the random year effects so that interannual variation is driven solely by climate. We fit the random year effects in the vital rate regressions to avoid over-attributing variation to climate covariates. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size _v_ at time $t+1$ is

@@

Since the IPM is spatially-implicit, we cannot calculate neighborhood crowding for specific genets ($w_{ij}$). Instead, we use an approximation ($\bar{\bold{w_{j}}}$) that captures the essential features of neighborhood interactions [@Adler2010]. We also use a 'no-overlap' rule for conspecific genets to approximate the overdispersion of large genets in space [@Adler2010; @Adler2012; @Chu2015].

@@

which means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species. Our stastical model for recruitment (_R_, described above) returns the number of new recruit produced per quadrat. Following previous work [@Adler2012; @Chu2015], we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eqs. 9-10). We use the same approach for drawing parameter values as described for the IPM. After drawing the appropriate parameter set, we calculate the mean response (population cover at *t+1* = $x_{t+1}$) according to Eq. 9. We then make a random draw from a [0,1] truncated lognormal distribution with mean equal to $x_{t+1}$ from Eq. 9 and the variance estimate from the fitted model. We can then project the model forward by drawing a new parameter set (unique to climate year and MCMC iteration) at each timestep. As with the IPM, random year effects are ignored for all simulations.

@@

### Testing sensitivity to climate covariates
To see if models based on aggregated data are as sensitive to climate covariates as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

Our expectation is that the IPM will produce more accurate and precise forecasts. This could be due to (1) the smaller  sample size of the quadrat level data sets compared to the individual level data sets, leading to larger parameter uncertainty for the QBM, or (2) the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty we refit the QBM statistical model (Eqs. 9-10) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptLag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress sample size against parameter uncertainty.

@@

Results
-------
### Comparison of forecast models
The IPM had significantly lower overall error (MAE, mean absolute error) for three species (*B. gracilis*, *H. comata*, *P. smithii*; Table 1). In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

@@

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6; simulations with mean parameters are in Appendix D for comparison). As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to a very high instrinsic rate of recruitment combined with uncertainty in climate coefficients which lead to high recruitment boom years and subsequent busts when young plants suffer high mortality (Appendx C). When we included model error and parameter uncertainty, forecast changes in proportional cover always spanned a wide range of negative to positive values. In other words, neither model could predict whether a climate perturbation would increase or decrease equilibrium population size.

@@

```{r comment_out, echo=FALSE, eval=FALSE}
# Equilibrium cover from both models was sensitive to climate (Fig. 3a-d). The IPM projected percent changes in equilibrium cover from `r round(min(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)`% for *B. gracilis*, `r round(min(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)`% for *H. comata*, `r round(min(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)`% for *P. smithii*, and `r round(min(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)`% for *P. secunda*. The QBM projected opposite and greater percent changes in equilibrium cover for *B. gracilis* (`r round(min(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)`%) and *H. comata* (`r round(min(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)`%; Fig. 3a-b). For *P. smithii*, the QBM projected opposite changes in equilibrium cover than the IPM, but of similar magnitude (`r round(min(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)`%; Fig. 3c). *P. secunda* was the only species that the IPM and QBM made projections of the same sign and somewhat similar magnitude (`r round(min(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)`%; Fig. 3d). 
```

@@

Our comparison of a traditional, demographic population model (the IPM) with a model inspired by density-structured models (the QBM) showed that the IPM outperformed the QBM: the IPM was more accurate and precise than the QBM in out-of-sample cross validation (Table 1). The superiority of the IPM could reflect either differences in sample size or the effect of averaging over unique effects of climate on each individual-level vital rate. Although increasing sample size of quadrat percent-cover observations would be easy to do in the field, we found little evidence that it would lead to higher precision of climate coefficient estimates (Fig. 4). We did, however, find evidence that the QBM statistical model failed to identify climate dependence for some vital rates (Fig. 5). For no species were climate effects from the QBM strongly correlated with all three vital rates (Fig. 5). @Freckleton2011 acknowledge that averaging over complex stage dependence will lead to poorly specified models. This is analagous to our situation, but instead of averaging over complex life histories, we are averaging over complex climate dependence.

@@

An important, but unexpected, result of our analysis was the great uncertainty in forecasts, even for our best model. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model, usually allowing for some variability around the determinstic process [e.g. @Battin2007; @Jenouvrier2009; @Adler2012]. If we follow tradition and calculate the mean response to climate perturbation with only model error and interannual variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). It would be tempting to interpret this inconsistency as further evidence for the superiority of the IPM. However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6), though the IPM projections are generally more precise (consistent with our cross-validation results). The real story is that both models produce highly uncertain forecasts. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. \todo{Add in sentence on larger perturbations.}

@@

Our results also offer a cautionary tale because forecast uncertainty was large for both model types. Even with 14 years of detailed demographic data and sophisticated modeling techniques, we failed to produce forecasts with any level of acceptable uncertainty. Uncertainty of climate change related forecasts can be reduced by (1) longer time series and (2) climate covariates that match the scale of inference (e.g., plot rather than landscape level climate/weather metrics).

@@

```{r figure_6, dependson="plot-options", fig.cap="Equilibrium cover and 90% quantiles around the mean prediction when model error and parameter uncertainty are propogated through the simulation phase. Climate simulations are as in Figure 3.", fig.width=4.5, fig.height=7, cache=FALSE, results='hide'}
####
##  Vary parameter runs
####
setwd("~/Repos/MicroMesoForecast/analysis/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/climchange_varyparams/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
torm <- c(grep("10", ipmfiles_mean), grep("20", ipmfiles_mean), grep("30", ipmfiles_mean))
ipmfiles_mean <- ipmfiles_mean[-torm]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]
